---
title: "Creation of ADaM datasets Subject-Level Analysis (ADSL) "
author: "Phanikumar s tata"
date: "9/6/2020"
output: html_document
mainfornt: Arial
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#install.packages("devtools")
#library(devtools)
#install_github("sas2r/clinical_fd")
#install.packages("naniar")              
#install.packages("tidyverse")
#install.packages("htmltools")
#install.packages("reshape2")
library(clinicalfd)
library(dplyr)
library(knitr)
library(naniar)

#install.packages("diffdf")
 library(diffdf)
 
 #install.packages("arsenal")
 library(arsenal)
 
library(reshape2)
library(gtsummary)

```

```{r,echo=TRUE}
# knitr::kable(names(dm) , format = "markdown")
# names(dm)
# knitr::kable(head (dm))
```

#Content of SDTM.DM dataset 

\pagebreak
#Get the names of SDTM.dm data and ADS.adsl 

# Variables present in SDTM.DM
1. "studyid"  "domain"   "usubjid"  "subjid"   "rfstdtc"  "rfendtc"  "rfxstdtc" \line
2. "rfxendtc" "rficdtc"  "rfpendtc" "dthdtc"   "dthfl"    "siteid"   "age"     \line
3. "ageu"     "sex"      "race"     "ethnic"   "armcd"    "arm"      "actarmcd" \line
4. "actarm"   "country"  "dmdtc"    "dmdy"    \line

\pagebreak
 
# Variables needed in ADSL \line 
 1. "studyid"  "usubjid"  "subjid"   "siteid"   "sitegr1"  "arm"      "trt01p"     \line
 2. "trt01pn"  "trt01a"   "trt01an"  "trtsdt"   "trtedt"   "trtdur"   "avgdd"     \line
 3. "cumdose"  "age"      "agegr1"   "agegr1n"  "ageu"     "race"     "racen"     \line
 4. "sex"      "ethnic"   "saffl"    "ittfl"    "efffl"    "comp8fl"  "comp16fl"  \line
 5. "comp24fl" "disconfl" "dsraefl"  "dthfl"    "bmibl"    "bmiblgr1" "heightbl"  \line
 6. "weightbl" "educlvl"  "disonsdt" "durdis"   "durdsgr1" "visit1dt" "rfstdtc"   \line
 7. "rfendtc"  "visnumen" "rfendt"   "dcdecod"  "dcreascd" "mmsetot"              \line
 
\pagebreak
 
# Derivation rule Subject-Level Analysis (ADSL) adsl.xpt

1.STUDYID Study Identifier text 12   DM.STUDYID \line

2.USUBJID Unique Subject Identifier text 11   DM.USUBJID\line

3.SUBJID Subject Identifier for the Study text 4   DM.SUBJID \line

4.SITEID Study Site Identifier text 3   DM.SITEID \line

5.SITEGR1 Pooled Site Group 1 text 3   refer to SAP, Section 7.1 - if not pooled then                   SITEGR1=SITEID. If pooled, SITEGR1 will be 900  \line

6.ARM     Description of Planned Arm text 20  ARM DM.ARM  \line 

7.TRT01P  Planned Treatment for Period 01 text 20  ARM DM.ARM  \line

8.TRT01PN Planned Treatment for Period 01 (N) integer 8  ARMN Numeric code for TRT01P  
   which corresponds to the randomized dose \line

Code List - ARMN, Reference Name (ARMN)  

    Coded Value     Decode
    0               Placebo 
    54              Xanomeline Low Dose 
    81              Xanomeline High Dose    
   
9.TRT01A  Actual Treatment for Period 01 text 20  ARM TRT01A=TRT01P, i.e., no                   difference between actual and randomized treatment in this study.\line 

10.TRT01AN Actual Treatment for Period 01 (N) integer 8  ARMN Numeric code for TRT01A            which corresponds to the randomized dose \line

11.TRTSDT   Date of First Exposure to Treatment integer 8 DATE9.  SV.SVSTDTC when SV.VISITNUM=3, converted to SAS date \line


```{r,echo=FALSE}

# Excluding the Screen failure subjects

# Looks lot useful dput(head(dm))

#dm_01  <- dplyr:: filter(dm ,arm !="Screen Failure" )

#dm <- dplyr::mutate (dm_01 , usubjid = as.factor(usubjid))
#svstdtc <- select ( sv , usubjid  , svstdtc , visitnum )
#svstdtc <- filter (svstdtc , visitnum == 3.0 )

# Doing an inner join to get svstdtc filed 
#dm_01 <- dplyr::inner_join( dm_01 , svstdtc , by = "usubjid")

# When Joining two data frames by factors you get a note "Column `usubjid` joining factors with different levels, coercing to character vector" indicating that one of the columns which are being used in your join is clearly a factor and also that factor had clearly the different level in your dataset. 

# Note so if you do not want to loose any of your information then your factors should be converted to character values 


dm_01  <- dplyr:: filter(dm ,arm !="Screen Failure" )

# Checking the nature of the variables 
str(select(dm, usubjid )) 
str(select(sv, usubjid )) 

svstdtc <- select ( sv , usubjid  , svstdtc , visitnum )
svstdtc <- dplyr::mutate (svstdtc , usubjid = as.character(usubjid))
svstdtc <- filter (svstdtc , visitnum == 3.0 )

# Doing an inner join to get svstdtc filed
dm_01 <- mutate (dm_01 , usubjid = as.character(usubjid) )
dm_01 <- dplyr::inner_join( dm_01 , svstdtc , by = "usubjid")


#summary(dm_01)
#summary(adsl)

adsl_01 <- dplyr::select(dm_01 , studyid, usubjid , subjid , siteid , arm)

trt01pn <- dplyr::distinct(adsl, trt01pn , trt01p)

#using the Mutate
adsl_01 <- dplyr :: mutate( adsl_01 , sitegr1 = siteid, trt01p = arm ,TRT01A = arm )

# ifelse function gives you only True or False i.e 0 and 1 only 

adsl_01 <- adsl_01 %>% 
mutate( trt01pn  =ifelse( (arm %in% "Placebo" ) , 0 ,
                  ifelse( (arm %in% "Xanomeline Low Dose" ) , 54,
                  ifelse( (arm %in% "Xanomeline High Dose" ) , 81 , NA)))
        
        ) 

adsl_01 <- dplyr :: mutate( adsl_01 ,trt01an = trt01pn )




#Multiple if Else statements can be written similarly to excel's if function
#Base R syntax

adsl_01$trt01pn_new = ifelse ( adsl_01$arm %in% "Placebo" , 0  ,
                      ifelse ( adsl_01$arm %in% "Xanomeline Low Dose"  ,  54 ,
                      ifelse ( adsl_01$arm %in% "Xanomeline High Dose"  ,  81, NA ))) 

# The thing to be observed below is that flag value is difined both numeric and character 1 and Diff , R won't complain 

adsl_01 <- adsl_01 %>% 
           mutate( compare_flag = ifelse ( trt01pn_new != trt01pn , "Diff" , 1 ))

adsl_01 <- adsl_01 %>% 
           mutate( compare_flag = ifelse ( trt01pn_new != trt01pn , "Diff" , "Same" ))

# NOTE using singe = rather than == as

#select (-arm)    '-' -ve sign for dropping the Arm variable as we are deriving trt01p 
adsl_01 <- dplyr::select(adsl_01 , - arm )

head(adsl_01)

```

# Here all the above steps are done in single step using "%>%" pipe operator . Also see the new derived variable sitegr1 is used in the select statement 

```{r label = "all_code"}

svstdtc <- sv %>% 
           filter(visitnum == 3.0) %>% 
           mutate (usubjid = as.character(usubjid)) %>% 
           select (usubjid , svstdtc)

dm_02  <- dplyr:: filter(dm ,arm !="Screen Failure" )
dm_02  <- dplyr:: mutate (dm ,usubjid = as.character(usubjid) )
# Doing an inner join to get svstdtc filed 


dm_02  <- dplyr::inner_join( dm_02 , svstdtc , by = "usubjid")


adsl_02 <- dm_02 %>% 
  filter (actarmcd !="Scrnfail") %>% 
  mutate ( 
    sitegr1 = siteid , trt01p = arm , trt01a = arm, 
    
    trt01pn  =ifelse( (arm %in% "Placebo" ) , 0 ,
                    ifelse( (arm %in% "Xanomeline Low Dose" ) , 54,
                    ifelse( (arm %in% "Xanomeline High Dose" ) , 81 , NA)))  ,
    trt01an = trt01pn ,
    trtsdt = svstdtc ,
    trtsdtn = as.Date(as.character (trtsdt)),
                
    ittfl   = ifelse (trt01pn %in% c(0,54,81) , "Y" , "N" )  , 
    
    agegr1  = case_when( age < 65  ~ "<65" , 
                              (65 <= age & age <= 80 ) ~ "65-80"  , 
                               (age > 80 ) ~ ">80"  ),
           
    agegr1n  = case_when( age < 65  ~ 1  , 
                                (65 <= age & age <= 80 ) ~ 2  , 
                                (age > 80 ) ~ 3  ),
           
    race  = case_when( race %in% "WHITE" ~ "WHITE" , 
                              race %in% "BLACK OR AFRICAN AMERICAN" ~ "BLACK OR AFRICAN AMERICAN" ,
                              race %in% "AMERICAN INDIAN OR ALASKA NATIVE" ~ "AMERICAN INDIAN OR ALASKA NATIVE" ,
                              race %in% "ASIAN" ~ "ASIAN" 
                              ) , 
    racen  = case_when( race %in% "WHITE" ~ 1 , 
                              race %in% "BLACK OR AFRICAN AMERICAN" ~ 2 ,
                              race %in% "AMERICAN INDIAN OR ALASKA NATIVE" ~ 6 ,
                              race %in% "ASIAN" ~ 7 ),
           
    sex  =  case_when(sex %in% "F" ~ "Female"  , 
                             sex %in% "M" ~ "Male"    , 
                             sex %in% "U" ~  "Unknown" ) , 
           
    ethnic  =  case_when(ethnic %in% "NOT HISPANIC OR LATINO" ~ "NOT HISPANIC OR LATINO"  , 
                                ethnic %in% "HISPANIC OR LATINO" ~ "HISPANIC OR LATINO"    , 
                                ethnic %in% "U" ~  "Unknown" ),
    
    saffl =   case_when ((ittfl %in% "Y" & !is.na(trtsdtn))    ~ "Y" ) 
    
  ) %>% 
  
  select ( studyid , usubjid , subjid , siteid , sitegr1 , trt01p ,
           trt01pn , trt01a , trt01an , - svstdtc , trtsdt , age , agegr1 , 
           agegr1n , race ,racen , sex , ethnic  , ittfl , trt01pn, trtsdtn , saffl)

#head(adsl_02)

```
 
12.TRTEDT Date of Last Exposure to Treatment integer 8 DATE9.  The date of final dose (from the CRF) is EX.EXENDTC on the subject's last EX record. If the date of final dose is missing for the subject and the subject discontinued after visit 3, use the date of discontinuation as the date of last dose. Convert the date to a SAS date.  \line

13. TRTDUR Duration of Treatment (days) integer 8   TRTEDT-TRTSDT+1 \line

```{r label = "TRTEDT Derivation"}
ex_01 <- arrange (select (ex , usubjid , exseq , extrt , visit, visitnum , exendtc) , 
                  usubjid , visitnum ) 
  
ex_02 <- ex_01 %>% 
                arrange (usubjid , visitnum) %>% 
                group_by (usubjid ) %>% 
                summarise_all (last) %>% 
                mutate (trtedt = exendtc)
 
 
#Getting the records with missing dates  
ex_02_missing_date <-  filter (ex_02, trtedt < 0   ) 

#Getting the records with non-missing dates 
ex_03 <- filter (ex_02 , trtedt >0 )

ex_03 <- select (ex_03 , usubjid , trtedt )

ds_disp <- filter ( ds, dscat == "DISPOSITION EVENT") 


ex_02_missing_date  <- dplyr::mutate (ex_02_missing_date ,usubjid = as.character(usubjid) )
ds_disp  <- dplyr:: mutate (ds_disp ,usubjid = as.character(usubjid) )
 

ds_disp2  <- dplyr::inner_join( select(ex_02_missing_date,usubjid)  ,
                                select (ds_disp , usubjid ,dsstdtc ) ,
                                        by = "usubjid")
ds_disp2 <-  ds_disp2 %>% 
               mutate ( trtedt = dsstdtc) %>% 
               select(usubjid , trtedt)

trtedt <-  rbind (ex_03  , ds_disp2)

#Getting it back to main dataset
adsl_02  <- dplyr::inner_join( adsl_02 ,
                               trtedt ,
                               by = "usubjid")
 
#Numerical dates trtsdtn trtedtn for caliculations 
#TRTDUR Duration of Treatment (days) integer 8   TRTEDT-TRTSDT+1

adsl_02 <- adsl_02 %>% 
                 mutate (trtsdtn = as.Date(as.character (trtsdt)) ,
                         trtedtn = as.Date(trtedt)   ,
                         trtdur = (trtedtn - trtsdtn ) + 1  
                         ) 

# For Comparing you do not need to create a new variable as R creates x. and y. variable
#in the joing table 

# adsl      <- adsl %>% 
#              mutate(trtedt_prod = trtedt) 
# 
# 
# adsl_trtedt_chk <- dplyr::inner_join( adsl ,
#                                       trtedt ,
#                                       by = "usubjid")
#  
``` 
EFFFL Efficacy Population Flag  text 1 ["N" = "No", "Y" = "Yes"]  <YN>
  
Y if SAFFL='Y AND subject has at least one record in 
QS for ADAS-Cog with VISITNUM>3 AND 
at least one record in QS for CIBIC+ with VISITNUM>3, N otherwise 

```{r}
# TIP Rename works left to right meaning 
# new name is on the left while old name is right 
qs_efffl_all <- qs %>% 
  filter( (qstestcd %in% c("ACTOT" , "CIBIC")  & visitnum >3)  )

qs_efffl01 <- qs %>% 
            filter( (qstestcd %in% c("ACTOT")  & visitnum >3)  ) %>% 
            rename (actot = qstestcd , actot_v = visitnum) %>% 
            select(usubjid , actot_v , actot)
 
qs_efffl01 <- dplyr::distinct(qs_efffl01, usubjid , .keep_all = TRUE )


qs_efffl02 <- qs %>% 
  filter( (qstestcd %in% c( "CIBIC")  & visitnum >3)  ) %>% 
  rename ( cibic = qstestcd , cibic_v = visitnum) %>% 
  select(usubjid , cibic_v , cibic)

qs_efffl02 <- dplyr::distinct(qs_efffl02, usubjid , .keep_all = TRUE )

qs_all <- full_join( qs_efffl01 , qs_efffl02 , by = "usubjid" ) 

#   subject has at least one record in QS for ADAS-Cog with VISITNUM>3
#   AND at least one record in QS for CIBIC+ with VISITNUM>3, N otherwise

qs_all <- qs_all %>% 
          filter (  !is.na(actot) & !is.na(cibic) ) %>% 
          mutate ( qs_flag = 1 , usubjid = as.character(usubjid))

 
adsl_02 <-left_join( adsl_02 , qs_all , by = "usubjid" )
 
adsl_02  <- adsl_02 %>%
           mutate ( qs_flag = coalesce(qs_flag , 0 ) , 
               efffl  = case_when ((qs_flag == 1 & saffl == "Y")  ~ "Y" ,
                                   (qs_flag == 0 & saffl == "Y" )  ~ "N" )) 


```

COMP8FL Completers of Week 8 Population Flag  text 1 ["N" = "No", "Y" = "Yes"]
Y if subject has a SV.VISITNUM=8 and ENDDT >= date of visit 8, N otherwise 

COMP16FL Completers of Week 16 Population Flag  text 1 ["N" = "No", "Y" = "Yes"]
Y if subject has a SV.VISITNUM=10 and ENDDT>=date of visit 10, N otherwise 

COMP24FL Completers of Week 24 Population Flag  text 1 ["N" = "No", "Y" = "Yes"]
Y if subject has a SV.VISITNUM=12 and ENDDT>= date of visit 12 , N otherwise 

```{r}
# adsl_COMP8FL  <- select (adsl , usubjid   , comp8fl ,comp16fl , trtedt  ) %>% 
#                    mutate(trtedt_p = trtedt ) %>% 
#                    select ( - trtedt )

svstdtc <- select ( sv , usubjid  , svstdtc , visitnum  )

sv_visits <- sv %>% 
  mutate ( usubjid = as.character(usubjid), 
           svstdtn = as.Date(as.character(svstdtc)  )) %>% 
  select ( usubjid  , svstdtc , visitnum , visit )


sv_visit08 <-  sv_visits %>%  filter (visitnum == 8.0 ) %>% 
              mutate(svstdtc_8 = as.Date(as.character(svstdtc) ) )  %>%
                        select ( usubjid , svstdtc_8 )

sv_visit10 <-  sv_visits %>%  filter (visitnum == 10.0 ) %>% 
                mutate(svstdtc_10 = as.Date(as.character(svstdtc) ) )  %>%
                select ( usubjid , svstdtc_10 )

#For Comparing purpose using wrong visitnum , should be 
# visitnum == 12.0 

sv_visit12 <-  sv_visits %>%  filter (visitnum == 12.0 ) %>% 
               mutate(svstdtc_12 = as.Date(as.character(svstdtc) ) )  %>%
               select ( usubjid , svstdtc_12 )

adsl_trt_edtn  <- select ( adsl_02 , usubjid , trtedtn )

adsl_02_comp  <-   adsl_trt_edtn  %>%
              full_join( sv_visit08      , by = "usubjid") %>% 
              full_join(sv_visit10       , by = "usubjid") %>% 
              full_join(sv_visit12       , by = "usubjid") %>% 
  
              mutate( comp8fl = ifelse (trtedtn >= svstdtc_8 , "Y" , "N" ),
                      comp8fl = coalesce(comp8fl  , "N" ) , 
                      
                      comp16fl = ifelse (trtedtn >= svstdtc_10 , "Y" , "N" ),
                      comp16fl = coalesce(comp16fl  , "N" ) , 
                      
                      comp24fl = ifelse (trtedtn >= svstdtc_12 , "Y" , "N" ),
                      comp24fl = coalesce(comp24fl  , "N" )
                      
                      )  %>% 
                     select (-trtedtn)
  

# Merging it back to main dataset 

adsl_02   <-   adsl_02  %>%
              full_join( adsl_02_comp      , by = "usubjid") 



  
# 
# adsl_02_comp  <- sv_visit08 %>%
#               full_join(adsl_02       , by = "usubjid") %>% 
#               full_join(sv_visit10    , by = "usubjid") %>% 
#               full_join(adsl_COMP8FL  , by = "usubjid") %>% 
#   
#               mutate( comp8fl_w = ifelse (trtedtn >= svstdtc_8 , "Y" , "N" ),
#                       comp8fl_w = coalesce(comp8fl_w , "N" ) , 
#                       
#                       comp16fl_w = ifelse (trtedtn >= svstdtc_10 , "Y" , "N" ),
#                       comp16fl_w = coalesce(comp16fl_w , "N" ) 
#                       
#                       ) %>% 
#               select (usubjid , comp16fl ,comp16fl_w  ,trtedt_p, trtedtn , svstdtc_10 )
```

DISCONFL Did the Subject Discontinue the Study?  text 1 ["Y" = "Yes"]
Y if DCREASCD ^= 'Completed'. Null otherwise

DSRAEFL Discontinued due to AE?  text 1 ["Y" = "Yes"]
Y if DCREASCD='Adverse Event'. Null otherwise
 
DTHFL Subject Died?  text 1 ["Y" = "Yes"]
DM.DTHFL 

```{r}

# DISCONFL Did the Subject Discontinue the Study?  text 1 ["Y" = "Yes"]
# Y if DCREASCD ^= 'Completed'. Null otherwise

disconfl <- ds %>% 
            filter ( dsdecod == "COMPLETED") %>% 
            mutate( disconfl = "Y", usubjid = as.character(usubjid)) %>% 
            select ( usubjid , disconfl)

# DSRAEFL Discontinued due to AE?  text 1 ["Y" = "Yes"]
# Y if DCREASCD='Adverse Event'. Null otherwise

dsraefl  <- ds %>% 
  filter ( dsdecod == "ADVERSE EVENT") %>% 
  mutate( dsraefl = "Y" , usubjid = as.character(usubjid)) %>% 
  select ( usubjid , dsraefl)

# DTHFL Subject Died?  text 1 ["Y" = "Yes"]
# DM.DTHFL 

dthfl02   <- dm %>% 
  filter (  dthfl == "Y") %>% 
  mutate( usubjid = as.character(usubjid) ) %>% 
  select (  usubjid , dthfl )
  

dthfl   <- ds %>% 
  filter ( dsdecod == "DEATH") %>% 
  mutate( dthfl = "Y" , usubjid = as.character(usubjid) ) %>% 
  select ( usubjid , dthfl)


adsl_02 <- mutate (adsl_02 , usubjid = as.character(usubjid))

adsl_02  <-   adsl_02  %>%
                  full_join( disconfl         , by = "usubjid") %>% 
                  full_join( dsraefl          , by = "usubjid") %>% 
                  full_join( dthfl02            , by = "usubjid") 
                  
```
HEIGHTBL Baseline Height (cm)  float 8  Derived: 
VSSTRESN when VS.VSTESTCD='HEIGHT' and VS.VISITNUM=1
 
WEIGHTBL Baseline Weight (kg)  float 8  Derived: 
VSSTRESN when VS.VSTESTCD='WEIGHT' and VS.VISITNUM=3 

BMIBL Baseline BMI (kg/m^2)  float 8  Derived: 
WEIGHTBL / ((HEIGHTBL*100)**2)
 
BMIBLGR1 Pooled Baseline BMI Group 1  text 6 ["<25" = "Normal", "25-<30" = "Overweight", ">=30" = "Obese"]

BMIBLGR1=Normal if . < BMIBL <25. BMIBLGR1=Overweight if 25 <=BMIBL <30. BMIBLGR1=Obese if BMIBL >=30


```{r}

# HEIGHTBL Baseline Height (cm)  float 8  Derived: 
#   VSSTRESN when VS.VSTESTCD='HEIGHT' and VS.VISITNUM=1

heightbl <- vs %>% 
            filter (vstestcd == "HEIGHT" & visitnum == 1 ) %>% 
            mutate ( heightbl = vsstresn ) %>% 
            select ( usubjid , heightbl )%>% 
          mutate (usubjid = as.character(usubjid))

# WEIGHTBL Baseline Weight (kg)  float 8  Derived: 
#   VSSTRESN when VS.VSTESTCD='WEIGHT' and VS.VISITNUM=3 

weightbl <- vs %>% 
  filter (vstestcd == "WEIGHT" & visitnum == 3) %>% 
         rename (weightbl = vsstresn ) %>% 
          select ( usubjid , weightbl ) %>% 
          mutate (usubjid = as.character(usubjid))

# BMIBL Baseline BMI (kg/m^2)  float 8  Derived: 
# WEIGHTBL / ((HEIGHTBL*100)**2) 
# Check this derivation *100 

heightbl <- mutate (heightbl , usubjid = as.character(usubjid))
weightbl <- mutate (weightbl , usubjid = as.character(usubjid))


bmi  <-  inner_join(heightbl , weightbl , by = "usubjid" ) 

bmi <- bmi %>% 
       mutate ( bmi = ((weightbl / (heightbl ^ 2 ) *10000) )  ) 

bmi <- mutate (bmi , usubjid = as.character(usubjid))


adsl_02 <- full_join ( adsl_02 , bmi , by = "usubjid")

```
 



# CUMDOSE Cumulative Dose (as planned) float 8
# For ARMN=0 or 1: CUMDOSE=TRT01PN*TRTDUR. ---

```{r}
# If treatment arm in 0 or 54 then it is just trt01pn* trtdur
cumdose_01  <- adsl_02 %>% 
           filter( ( trt01pn %in% c(0,54) ) ) %>% 
           mutate ( cumdose = trt01pn * trtdur ) %>% 
           select ( usubjid , trt01pn , trtdur , cumdose ) 

```


```{r}

svstdtc <- select ( sv , usubjid  , svstdtc , visitnum  )

sv_visits <- svstdtc %>% 
  mutate ( usubjid = as.character(usubjid), 
           svstdtn = as.Date(as.character(svstdtc)  ))

sv_visit03 <-  sv_visits %>% 
  filter (visitnum %in% 3.0 ) %>% 
  mutate(svstdtn_3 = svstdtn ) %>% 
  select ( -visitnum , -svstdtc , -svstdtn)

sv_visit04 <-  sv_visits %>% 
  filter (visitnum %in% 4.0 ) %>% 
  mutate(svstdtn_4 = svstdtn ) %>% 
  select ( -visitnum , -svstdtc , -svstdtn)

sv_visit09 <-  sv_visits %>% 
  filter (visitnum %in% 9.0 ) %>% 
  mutate(svstdtn_9 = svstdtn ) %>% 
  select ( -visitnum , -svstdtc , -svstdtn)


# Doing an inner join to get svstdtc filed
#TIP when there are more than two dataframess
# Inner Join this way does not work 
#sv_visits <- dplyr::inner_join( sv_visit03,
                                # sv_visit04,
                                # sv_visit09, 
                                # by = "usubjid")
#other way to do merge is 
# #filter (actarmcd !="Scrnfail") %>% 
#   mutate ( 
#     sitegr1 = siteid , trt01p = arm , trt01a = arm, 
#     trt01pn  =ifelse( (arm %in% "Placebo" ) , 0 ,
#                       ifelse( (arm %in% "Xanomeline Low Dose" ) , 54,
#                               ifelse( (arm %in% "Xanomeline High Dose" ) , 81 , NA)))  , 
#     

adsl_cumdose_03 <- select ( adsl , usubjid , trt01pn , trtdur , cumdose ,trtedt )
adsl_cumdose_03 <- dplyr:: mutate (adsl_cumdose_03 , usubjid = as.character(usubjid))

sv_visits  <- sv_visit03 %>%
               inner_join(adsl_cumdose_03 , by = "usubjid") %>% 
               inner_join(sv_visit04 , by = "usubjid") %>% 
               inner_join(sv_visit09 , by = "usubjid")

 
sv_visits02  <- sv_visits %>% 
                 
                mutate ( cumdose_01 = trt01pn * trtdur , 
                         trt_dur02 = trtedt - svstdtn_4 , 
                         cumdose_02 = trt_dur02 * trt01pn )   %>% 
  
                select (usubjid ,trtdur, trt01pn, svstdtn_3 , svstdtn_4 , svstdtn_9 ,
                        cumdose_01 ,  cumdose  , trt_dur02  , cumdose_02)
 
```

#Comparing the dataframes 

```{r}
adsl <- adsl
adsl_prog  <- adsl %>% 
              mutate ( comp8fl = as.character(comp8fl) , 
                       comp16fl = as.character(comp16fl) , 
                       comp24fl = as.character(comp24fl) , 
                       efffl = as.character(efffl)  , 
                       ittfl = as.character(ittfl) , 
                       trtedt  = as.character(trtedt) , 
                       agegr1n = as.numeric(agegr1n)
                       ) %>% 
                     select (  usubjid , trt01pn , ittfl ,trtedt, 
                       efffl, trtdur, age, agegr1n  ,comp8fl , comp16fl , comp24fl )

adsl_vali  <- dplyr::select ( adsl_02 , usubjid , trt01pn , trtedt, efffl, 
                       trtdur , ittfl , agegr1n, comp8fl , comp16fl , comp24fl )

#compare differences using Dplyr 

dplyr:: all_equal( adsl_prog , adsl_vali  , convert = TRUE  , 
                   , ignore_col_order = TRUE,
                   ignore_row_order = TRUE )

#dplyr:: setdiff( adsl_prog , adsl_vali)

library (janitor)

#compare differences using Janitor
 compare_df_cols( adsl_prog , adsl_vali , return = "mismatch")
 
 #install.packages("diffdf")
 #library(diffdf)
 
 diffdf(adsl_prog , adsl_vali) 
 
 #install.packages("arsenal")
 #library(arsenal)
 arsenal::comparedf ( adsl_prog , adsl_vali)

 summary(comparedf ( adsl_prog , adsl_vali))

```

#Cleaner variables proc Compare example to be presented for presentation 

```{r}

# adsl_prog  <- adsl %>% 
#               mutate ( comp8fl = as.character(comp8fl) , 
#                        comp16fl = as.character(comp16fl) , 
#                        comp24fl = as.character(comp24fl) , 
#                        efffl = as.character(efffl)  , 
#                        ittfl = as.character(ittfl) , 
#                        trtedt  = as.character(trtedt) , 
#                        agegr1n = as.numeric(agegr1n)
#                        ) %>% 
#                      select (  usubjid , trt01pn , ittfl ,trtedt, 
#                        efffl, trtdur, age, agegr1n  ,comp8fl , comp16fl , comp24fl )
# 
# adsl_vali  <- dplyr::select ( adsl_02 , usubjid , trt01pn , trtedt, efffl, 
#                        trtdur , ittfl , agegr1n, comp8fl , comp16fl , comp24fl )
adsl <- adsl

adsl_prog <- adsl %>% 
            mutate  (
                     usbjid = as.character(usubjid), 
                     ethnic = as.character(ethnic), 
                     race = as.character(race),
                     sex = as.character(sex) , 
                     efffl = as.character(efffl) , 
                     saffl = as.character(saffl)
                     ) %>% 
             select ( usubjid , race , sex , ethnic , saffl , efffl  )



adsl_vali <- adsl_02 %>% 
        select ( usubjid , race , sex , ethnic , saffl , efffl )

#compare differences using Dplyr 

# dplyr:: all_equal( adsl_prog , adsl_vali  , convert = TRUE  , 
#                    , ignore_col_order = TRUE,
#                    ignore_row_order = TRUE )
# 
# dplyr:: setdiff( adsl_prog , adsl_vali)
# 
# library (janitor)
# 
# #compare differences using Janitor
#  compare_df_cols( adsl_prog , adsl_vali , return = "mismatch")
 
 #install.packages("diffdf")
 library(diffdf)
 
 diffdf(adsl_prog , adsl_vali) 
 
 #install.packages("arsenal")
 library(arsenal)
 arsenal::comparedf ( adsl_prog , adsl_vali)

 summary(comparedf ( adsl_prog , adsl_vali))

```

```{r}
# library(knitr)
#   knitr::opts_chunk$set(tidy = y , 
#                         fig.width =  10 , 
#                         fig.height = 5 , 
#                         fig.align = 'left', 
#                         warning =  FALSE ,
#                         message = FALSE , 
#                         echo = TRUE )
options (width = 120 )


#custom function for std err of the mean 
 
se = function (x) {
      data = na.omit (x) 
      sd(data)/ sqrt(length(data))
}

bmi_by_trt <- adsl %>% 
              group_by (trt01a) %>% 
              summarise( 
                ct = n( ) , 
                mn = mean (bmibl) , 
                median = median(bmibl),
                min = min(bmibl) , 
                max = max(bmibl) , 
                range = max-min , 
                sd = sd (bmibl) , 
                se = se(bmibl) , 
                sum = sum(bmibl)
                ) %>% 
               ungroup()

# Kurtosis and skew are not working 
#Adding labels to columns 
colnames( bmi_by_trt) = 
  c('BMI' , 'Count' , 'Mean' , 'Median' , 
    'Min'  , 'Max' , 'Range', 'Std Dev' , 'Std Error' , 
     'Sum')

kable (bmi_by_trt , format = 'markdown', digits = 4 )
 

kable ( (bmi_by_trt_list <- adsl %>% 
             filter (trt01a %in% "Xanomeline Low Dose" & bmibl %in% NA  ) %>% 
             select (usubjid , trt01a , bmibl)) , 
       format = 'markdown',
       caption = 'Listing Xanomeline Low Dose subjects' ,
       digits =  4 )

dplyr::tibble(adsl)
str (dplyr::tibble(adsl))


```


```{r}
# R does not produce if any one of the observations is missing 
bmi_by_trt <- adsl %>% 
  filter ( !is.na (bmibl) ) %>% 
              group_by (trt01a) %>% 
              summarise( 
                ct = n( ) , 
                mn = mean (bmibl) , 
                median = median(bmibl),
                min = min(bmibl) , 
                max = max(bmibl) , 
                range = max-min , 
                sd = sd (bmibl) , 
                se = se(bmibl) , 
                sum = sum(bmibl)
                ) %>% 
              ungroup()

# Kurtosis and skew are not working 
#Adding labels to columns 
colnames( bmi_by_trt) = 
  c('Treatment' , 'Count' , 'Mean' , 'Median' , 
    'Min'  , 'Max' , 'Range', 'Std Dev' , 'Std Error' , 
     'Sum')

kable (bmi_by_trt , format = 'markdown', 
       caption = 'Summary statistics for BMI at baseline by Treatment' ,
       digits = 4 )
 
```

```{r}
#library(dplyr)

adsl2 <- adsl %>% 
         mutate ( trt01p = as.character(trt01p) , 
                  bmibl = as.numeric(bmibl) , 
                  heightbl = as.numeric(heightbl) , 
                  educlvl = as.numeric(educlvl) , 
                  ) %>% 
             select (trt01p, sex , race , bmibl, heightbl , educlvl) 
str(adsl2)
             
gtsummary::tbl_summary( adsl2 ,
                        by = trt01p   ) %>% 
                         add_overall()  %>% 
                         add_n ( ) %>% 
                         add_stat_label()  

adsl2 %>%
                tbl_summary(
                 # split the table by treatment variable
                 by = trt01p ,
                 type = list(c(sex,race) ~ "categorical") ,
                #change statistics printed in the table
                statistic = list(all_continuous() ~ "{mean} ({sd})") ,
                                  all_categorical() ~ "{n} / {N} ({p}%)"
                 )
```


```{r}
se = function (x) {
  data = na.omit (x) 
  sd(data)/ sqrt(length(data))
}

bmi_by_trt <- adsl %>% 
  filter ( !is.na (bmibl) ) %>% 
  group_by (trt01a) %>% 
  summarise( 
    n = n( ) , 
    mn = round(mean (bmibl) , digits = 2)  , 
    median = median(bmibl),
    min = min(bmibl) , 
    max = max(bmibl) , 
    range = max-min , 
    sd = round(sd (bmibl), digits = 2) , 
    se = round(se(bmibl) , digits = 2) , 
    sum = sum(bmibl)
  )

bmi_by_trt <- bmi_by_trt %>% 
              mutate (mean_sd  = paste0( mn ,"(" , sd  , ")" )) %>% 
              select( trt01a, n ,sum ,mean_sd, min , median , max , range , se )

   
bmi_by_trt2 <- reshape2::melt(bmi_by_trt , 
                    id.vars = c("trt01a" ))


bmi_by_trt3 <- bind_cols( 
                  (bmi_by_trt2 %>% 
                     filter (trt01a == "Placebo") %>% 
                     select (variable , trt01a, value, -trt01a ) %>% 
                     rename (placebo = value ) ) ,
                  
                  (bmi_by_trt2 %>% 
                     filter (trt01a == "Xanomeline Low Dose") %>% 
                     select ( trt01a, value, -trt01a ) %>% 
                     rename (lowdose = value ) )  , 
                  
                  (bmi_by_trt2 %>% 
                     filter (trt01a == "Xanomeline High Dose") %>% 
                     select ( trt01a, value, -trt01a ) %>% 
                     rename (highdose = value ) )          
                         )


bmi_by_trt_frq <- adsl %>% 
  filter ( !is.na (sex) ) %>% 
  group_by (sex , trt01a) %>% 
  summarise(sex_count = n( ) ) %>% 
  mutate (sex_freq = paste0 (sex_count, " (" , round((sex_count/ sum(sex_count))*100 , digits = 2  ) , "%)"  ) ,  
          variable = sex )  %>% 
   select ( sex, trt01a , sex_freq ) %>% 
   arrange( trt01a )
 
bmi_by_trt_frq2 <- bind_cols( 
  (bmi_by_trt_frq %>% 
     filter (trt01a == "Placebo") %>% 
     select (sex , trt01a, sex_freq, -trt01a ) %>% 
     rename (placebo = sex_freq ) ) ,
  
  (bmi_by_trt_frq %>% 
     filter (trt01a == "Xanomeline Low Dose") %>% 
     select ( trt01a, sex_freq, -trt01a, -sex ) %>% 
     rename (lowdose = sex_freq , , ldose = sex ) )  , 
  
  (bmi_by_trt_frq %>% 
     filter (trt01a == "Xanomeline High Dose") %>% 
     select ( trt01a, sex_freq, -trt01a , -sex ) %>% 
     rename (highdose = sex_freq , hdose = sex ) )             
)


bmi_by_trt_frq2 <- dplyr:: rename( bmi_by_trt_frq2,   variable = sex)
bmi_by_trt_frq2 <- select (bmi_by_trt_frq2 , -hdose , -ldose)


temp_data <- data.frame( matrix(ncol= 3 , nrow = 1 ))          
c <- c("variable")          
colnames(temp_data) <- c 

temp_data <- data.frame(matrix(ncol= 4 , nrow = 1 ,
  dimnames = list(NULL, c("variable" , "placebo", "lowdose", "highdose"
 )) ))


temp_data0 <- dplyr:: mutate( temp_data,   variable = "Sex" , variable = as.character(variable)
                              , placebo = as.character(placebo)
                              , lowdose = as.character(lowdose)
                              , highdose = as.character(highdose)
                              )
temp_data1 <- dplyr:: mutate( temp_data,   variable = "Baseline BMI" , variable = as.character(variable)
                              , placebo = as.character(placebo)
                              , lowdose = as.character(lowdose)
                              , highdose = as.character(highdose))

# 
# temp_data1<- dplyr::tibble(bmi_by_trt3)
# 
# bmi_by_trt_frq2 <- dplyr::tibble(bmi_by_trt_frq2)

bmi_by_trt3 <- dplyr:: mutate ( bmi_by_trt3 , variable = as.character(variable)
                                , placebo = as.character(placebo)
                                , lowdose = as.character(lowdose)
                                , highdose = as.character(highdose)) 
 
bmi_by_trt_frq2 <- dplyr:: mutate (  bmi_by_trt_frq2 , variable = as.character(variable)
                                     , placebo = as.character(placebo)
                                     , lowdose = as.character(lowdose)
                                     , highdose = as.character(highdose)) 

 
s_stats <- bind_rows (temp_data1, bmi_by_trt3,temp_data0 ,bmi_by_trt_frq2 )

s_stats

```



```{r}

# bmi_by_trt <- adsl %>% 
#               group_by(trt01pn) %>% 
#               summarise_at( c( "trt01pn" , "bmibl" , "educlvl" ) ,
#                                 mean, median ,
#                                  na.rm = TRUE) 
# if applied on a grouped tibble , these operations are not applied to grouping
# variables. The behavior depends on wheather the selection is implict 
# (all and if selections ) or explict (at selections) .
# 
# Grouping variables covered by explict selections in summarise_at() are always
# an error . Add -group_cols() to vars () selection to avoid this 

# 
# 
# trt01pn <- setdiff ( adsl , group_vars(trt01pn))
# 
# bmi_by_trt <- adsl %>% 
#   
#   summarise_at( nms ,   mean  )  

# setwd("C:/Users/tatav/OneDrive/Desktop/sas2r")
#getwd()

# my_logfile = "mY_logfile.txt"
# my_console_appender = console_appender(layout = default_log_layout())
# my_file_appender = file_appender( my_logfile , append = TRUE , 
#                                    layout =  default_log_layout())
# 
# my_logger <- log4r::logger(threshold = "INFO" , 
#                             appenders = list(my_console_appender ,                                                    my_file_appender))
# 
# log4r_info <- function(){
#   log4r::info(my_logger , "Info_message.")
# }
# 
# 
# log4r_error <- function(){
#   log4r::error(my_logger , "Error_message.")
# }
# 
# log4r_debug <- function(){
#   log4r::debug(my_logger , "Debug_message.")
# }
# 
# log4r_debug
# log4r_info
# log4r_error
# 
# readLines(my_logfile)

# 
```

